Stable individual differences in strategies within, but not between, visual search tasks

### 2018-11-30 Manuscript/manuscript.tex
Fixing some typos

\begin{abstract}
A striking range of individual differences has recently been reported in three different visual search tasks. These differences in performance can be attributed to strategy, that is, the efficiency with which participants control their search to complete the task quickly and accurately. Here we ask if an individual's strategy and performance in one search task is correlated with how they perform in the other two. We tested 64 observers in the three tasks mentioned above over two sessions. Even though the test-retest reliability of the tasks is high, an observer’s performance and strategy in one task did not reliably predict their behaviour in the other two. These results suggest search strategies are stable over time, but context-specific. To understand visual search we therefore need to account not only for differences between individuals, but also how individuals interact with the search task and context. These context-specific but stable individual differences in strategy can account for a substantial proportion of variability in search performance.
\end{abstract}

@@

A striking example of the effect of strategy is given by Boot and colleagues\cite{boot2006}. They asked participants to monitor a cluttered display for an object changing colour or suddenly appearing. Large individual differences were found with respect to the number of saccades participants made while monitoring the stimulus, and this was negatively correlated with detection performance. Proulx\cite{proulx2011} demonstrated that participants differed in how they prioritised visual features (colour or orientation) when searching through a set of objects for one with a particular colour-orientation combination. These differences in search strategy explained some of the variation in the extent to which participants were distracted by an irrelevant object of a unique size. However, participants were largely unaware of the search strategy they employed, and the majority were unable, or unwilling, to comply with instructions to employ a particular strategy.

Eye movement strategies have also been shown to be an important source of individual differences in visual search efficiency. Nowakowska, Clarke \& Hunt\cite{nowakowska2017} designed a simple search paradigm to discriminate between optimal\cite{najemnik-geisler2008} and stochastic \cite{clarke2016} search strategies. The task involved searching through arrays of line segments arranged such that those on one side of the display all had a very similar orientation ("homogeneous"), while those on the other side had higher variance ("heterogeneous", see Figure \ref{fig:exampleStimuli} for an example). This meant that targets appearing on the homogeneous side would be highly salient, and targets on the heterogeneous side  would be much harder to find. With these stimuli, the optimal eye movement strategy is to only search the heterogeneous half , as targets on the homogeneous side can be detected with peripheral vision. We will refer to this paradigm here as the Split-Half Line Segment task (SHLS). While some participants initially searched the displays near optimally, others carried out strategies counter to this, failing to even match the performance of the stochastic searcher. Furthermore, the degree to which they made saccades in line with the optimal search strategy was strongly correlated with their reaction times. A related version of this paradigm has been used in research investigating eye movement strategies in response to (simulated) hemianopia\cite{nowakowska2016,nowakowska2018}, with similar conclusions: the full spectrum of individual differences in strategy was observed, from near-optimal to the opposite of optimal. It is therefore not possible to conclude whether optimal or stochastic models better describe search without first explaining individual variability.

@@

Previous research has investigated the relationship between these behaviours to psychometric measures, but to date, these differences have not shown strong correlations with other attributes. In the ACVS paradigm, Irons \& Leber found no evidence of a correlation between the proportion of optimal choices made by observers and measures of visual working memory; trait impulsivity\cite{irons-leber2016}; novelty seeking; need for cognition; and intolerance of uncertainty\cite{irons-leber2018}. Similarly, the differences foraging behaviour are not accounted for by working memory or inhibitory control\cite{johannesson2017}.

@@

\subsection{Participants}
64 students from the University of Aberdeen took part in this study\footnote{data from an additional 11 participants was discarded due to being recorded with an inappropriate screen resolution.}. Participants were compensated for their time with either  course credit or \pounds 15. All participants gave informed consent. The study was approved by the University of Aberdeen Psychology Ethics Committee.

@@

\begin{figure}
\centering
\includegraphics[width=6.5cm]{../Scripts/scratch/cor_comparison.pdf}
\caption{The between- and within-task correlations for the three different search tasks. The bars indicate the $95\%$ confidence intervals for Pearson's correlation coefficient. Blue bars represent test-retest scores for each task for reaction times (\textit{rt}), optimality (\textit{opt}) or run length (\textit{rl}). Yellow bars indicate how well the strategy measures predict reaction times, while the red bars show that performance in one task is not a good indication of performance in another, either for reaction times of strategy.}
\label{fig:all_the_cor}
\end{figure}

>>>>

\begin{abstract}
A striking range of individual differences has recently been reported in three different visual search tasks. These differences in performance can be attributed to strategy, that is, the efficiency with which participants control their search to complete the task quickly and accurately. Here we ask if an individual's strategy and performance in one search task is correlated with how they perform in the other two. We tested 64 observers in the three tasks mentioned above over two sessions. Even though the test-retest reliability of the tasks is high, an observer's performance and strategy in one task did not reliably predict their behaviour in the other two. These results suggest search strategies are stable over time, but context-specific. To understand visual search we therefore need to account not only for differences between individuals, but also how individuals interact with the search task and context. These context-specific but stable individual differences in strategy can account for a substantial proportion of variability in search performance.
\end{abstract}

@@

A striking example of the effect of strategy is given by Boot and colleagues\cite{boot2006}. They asked participants to monitor a cluttered display for an object changing colour or suddenly appearing. Large individual differences were found with respect to the number of saccades participants made while monitoring the stimulus, and the number of saccades was negatively correlated with detection performance. Proulx\cite{proulx2011} demonstrated that participants differed in how they prioritised visual features (colour or orientation) when searching through a set of objects for one with a particular colour-orientation combination. These differences in search strategy explained some of the variation in the extent to which participants were distracted by an irrelevant object of a unique size. However, participants were largely unaware of the search strategy they employed, and the majority were unable, or unwilling, to comply with instructions to employ a particular strategy.

Eye movement strategies have also been shown to be an important source of individual differences in visual search efficiency. Nowakowska, Clarke \& Hunt\cite{nowakowska2017} designed a simple search paradigm to discriminate between optimal\cite{najemnik-geisler2008} and stochastic \cite{clarke2016} search strategies. The task involved searching through arrays of line segments arranged such that those on one side of the display all had a very similar orientation (homogeneous), while those on the other side had higher variance (heterogeneous). See Figure \ref{fig:exampleStimuli} for an example. This meant that targets appearing on the homogeneous side would be highly salient, and targets on the heterogeneous side  would be much harder to find. With these stimuli, the optimal eye movement strategy is to only search the heterogeneous half , as targets on the homogeneous side can be detected with peripheral vision. We will refer to this paradigm here as the Split-Half Line Segment task (SHLS). While some participants initially searched the displays near optimally, others carried out strategies counter to this, failing to even match the performance of the stochastic searcher. Furthermore, the degree to which they made saccades in line with the optimal search strategy was strongly correlated with their reaction times. A related version of this paradigm has been used in research investigating eye movement strategies in response to (simulated) hemianopia\cite{nowakowska2016,nowakowska2018}, with similar conclusions: the full spectrum of individual differences in strategy was observed, from near-optimal to the opposite of optimal. It is therefore not possible to conclude whether optimal or stochastic models better describe search without first explaining individual variability.

@@

Previous research has investigated the relationship between these behaviours to psychometric measures, but to date, these differences have not shown strong correlations with other attributes. In the ACVS paradigm, Irons \& Leber found no evidence of a correlation between the proportion of optimal choices made by observers and measures of visual working memory; trait impulsivity\cite{irons-leber2016}; novelty seeking; need for cognition; and intolerance of uncertainty\cite{irons-leber2018}. Similarly, the differences in foraging behaviour are not accounted for by working memory or inhibitory control\cite{johannesson2017}.

@@

\subsection{Participants}
64 students with normal or corrected-to-normal vision from the University of Aberdeen took part in this study\footnote{data from an additional 11 participants was discarded due to being recorded with an inappropriate screen resolution. Another participant was exluded due to colour blindness.}. Participants were compensated for their time with either  course credit or \pounds 15. All participants gave informed consent. The study was approved by the University of Aberdeen Psychology Ethics Committee.

@@

\begin{figure}
\centering
\includegraphics[width=6.5cm]{../Scripts/scratch/cor_comparison.pdf}
\caption{The between- and within-task correlations for the three different search tasks. The bars indicate the $95\%$ confidence intervals for Pearson's correlation coefficient. Blue bars represent test-retest scores for each task for reaction times (\textit{rt}), optimality (\textit{opt}) or run length (\textit{rl}). Yellow bars indicate how well the strategy measures predict reaction times, while the red bars show that performance in one task is not a good indication of performance in another, either for reaction times or strategy.}
\label{fig:all_the_cor}
\end{figure}
